---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Get started

This vignette demonstrates how to use `pc_meta_correlations` on Seurat data and visualize the results.

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(pcMetaCorrelations)
```

## Linear-model mode

The default analysis mode is `lm`, which fits a separate linear model for each metadata/PC pair.

## Seurat built-in example

```{r}
if (!requireNamespace("Seurat", quietly = TRUE)) {
  stop("Please install Seurat to run this example.")
}

data("pbmc_small", package = "SeuratObject")
seurat_obj <- pbmc_small
res <- pc_meta_correlations(seurat_obj, reduction = "pca", mode = "lm")
head(res)
```

## Plot the Seurat example

```{r}
plot_pc_meta_heatmap(res)
plot_top_pc_meta(res, n = 5, value = "effect_size")
```

## Richer metadata example

The `SeuratData` package provides larger example datasets. `celegans.embryo` includes more than 20 metadata columns, making it a strong example for metadata-PC association discovery.

```{r}
if (requireNamespace("SeuratData", quietly = TRUE)) {
  SeuratData::InstallData("celegans.embryo")
  data("celegans.embryo", package = "celegans.embryo.SeuratData")
  celegans.embryo <- Seurat::UpdateSeuratObject(celegans.embryo)
  if (!"pca" %in% names(celegans.embryo@reductions)) {
    if (!length(Seurat::VariableFeatures(celegans.embryo))) {
      celegans.embryo <- Seurat::FindVariableFeatures(celegans.embryo, verbose = FALSE)
    }
    celegans.embryo <- Seurat::ScaleData(celegans.embryo, features = Seurat::VariableFeatures(celegans.embryo), verbose = FALSE)
    celegans.embryo <- Seurat::RunPCA(celegans.embryo, verbose = FALSE)
  }

  res2 <- pc_meta_correlations(celegans.embryo, reduction = "pca", mode = "lm")
  head(res2)
  plot_pc_meta_heatmap(res2, top_n = 20, top_pcs = 8)
  plot_top_pc_meta(res2, n = 10, value = "effect_size")
}
```

## Interpreting results

- `statistic` is the model coefficient for numeric metadata or the F-statistic for categorical metadata.
- `effect_size` summarizes association strength.
- `direction` is positive/negative for numeric metadata and the highest-scoring category for categorical metadata.
